Iterative Monomorphisation

### 2024-03-08 iterative_monomorphisation.tex
edits with Tanguy

\item[\labelitemi] Given a formula \(\varphi \in \Phi\) and a symbol \(f\) occurring in \(\varphi\), the set \(P(\varphi, f)\) consists of all tuples of type arguments to which \(f\) is applied in \(\varphi\) and which contains a type variable. For example, if \(\sym{foldl}\langle \ty{nat}, \ty{list}(\alpha)\rangle\) occurs in \(\varphi\), then \((\ty{nat}, \ty{list}(\alpha)) \in P(\varphi, \sym{foldl}) \).
\end{enumerate}

@@

   \item For each formula \(\varphi \in \Phi\) and for each symbol \(f\) occurring in \(\varphi\):
   \begin{enumerate}
    \item[2.1.] For each tuple \((\tau_1, \dots, \tau_n) \in  M(f)\)        and each tuple \((\upsilon_1, \dots, \upsilon_n) \in P(\varphi, f)\),
     for each \(i\), match \(\upsilon_i\) against \(\tau_i\), yielding the substitution \(\sigma_i\) in case of success.

@@

\SetKwFunction{IterMono}{iterative\_monomorphisation}
\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\Fn(){\IterMono{\(\Phi\)}}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}\SetKw{To}{to}

   \SetKwData{Loop}{loop\_nb}\SetKwData{NewFormulae}{\(\Psi\)}
   \SetKwData{Problem}{\(\Phi\)}\SetKwData{AllSubst}{\(S^{\text{used}}\)}
   \SetKwData{Limit}{total\_new\_formulae}

   \SetKwFunction{Init}{initialisation}\SetKwFunction{FStep}{formula\_mono\_step}
   \SetKwFunction{TyVars}{type\_variables}\SetKwFunction{MonoSubst}{mono\_substs}

   \KwData{set of formulae \(\Phi\)}
   \KwResult{set of monomorphic formulae}

   \((M^{\text{old}}, P^{\text{old}}) \leftarrow (\emptyset, \emptyset)\)\;
   \((M^{\text{new}}, P^{\text{new}}) \leftarrow \Init{\Problem}\)\;

@@

   \For{\(i = 1\) \To \textcolor{ourblueviolet}{\Loop}}{
      \((M^{\text{next}}, P^{\text{next}})\leftarrow(\emptyset,\emptyset)\)\;
      \ForEach{\(\varphi \in \Problem\)} {
         \((M^{\Delta}, P^{\Delta}(\varphi), S^\Delta)\leftarrow \FStep(M^{\text{old}}, P^{\text{old}}(\varphi), M^{\text{new}}, P^{\text{new}}(\varphi))\)\;
         \(\AllSubst\leftarrow\AllSubst\cup S^\Delta\)\;
         \(M^{\text{next}}\leftarrow M^{\text{next}}\cup M^{\Delta}\)\;
         \(P^{\text{next}}(\varphi)\leftarrow P^{\text{next}}(\varphi)\cup P^{\Delta}(\varphi)\)\;
      }

      \((M^{\text{old}}, M^{\text{new}})\leftarrow (M^{\text{old}}\cup M^{\text{new}}, M^{\text{next}})\)\;
      \((P^{\text{old}}, P^{\text{new}})\leftarrow (P^{\text{old}}\cup P^{\text{new}}, P^{\text{next}})\)\;

@@

   \NewFormulae \(\leftarrow\emptyset\)\;
   \ForEach{\(\varphi\in\Phi\) polymorphic}{
      \(V\leftarrow \TyVars{\(\varphi\)}\)\;
      \(S^{\text{mono}}\leftarrow\MonoSubst{\AllSubst, \(\emptyset\), \(\{\}\), \(V\)}\)\;
      \ForEach{\(\sigma\in S^{\text{mono}}\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|\NewFormulae|<\Limit\)}{
            \textcolor{black}{
            \(\NewFormulae\leftarrow\NewFormulae\cup \{\sigma(\varphi)\}\)\;}
         }
         {\Return \NewFormulae}
      }}
   }
   \Return \NewFormulae\;

}
\end{algorithm}
\end{quote}
\caption{Pseudocode for iterative monomorphisation}
\label{iter_mono}
\end{figure}

@@

   %\LinesNumbered
   \SetKwFunction{SubstGen}{subst\_generation}\SetKwFunction{TyVars}{type\_variables}
   \SetKwFunction{FMonoStep}{formula\_mono\_step}

   \Fn(){\FMonoStep{\(M^{\text{old}}, M^{\text{new}}, P^{\text{old}}(\varphi), P^{\text{new}}(\varphi)\)}}{
   \SetKwData{Subst}{substs}\SetKwData{TypeVars}{type\_variables}
   \SetKwData{MaxMono}{max\_mono\_args}
   \SetKwData{MaxPoly}{max\_poly\_args}
   \SetKwData{UsedSubst}{\(S^{\text{used}}\)}

@@

   \KwData{
      \hspace{-2.7em}\begin{minipage}[t]{.8\textwidth}
      \begin{enumerate}
         \item[] old and new monomorphic type argument maps \(M^{\text{old}}, M^{\text{new}}\)
         \item[] old and new non-monomorphic type argument maps \(P^{\text{old}}(\varphi), P^{\text{new}}(\varphi)\)
      \end{enumerate}
      \end{minipage}
   }
   \KwResult{
      \hspace{-2.7em}\begin{minipage}[t]{.8\textwidth}
      \begin{enumerate}
         \item[] monomorphic type argument map
         \item[] non-monomorphic type argument map
         \item[] set of substitutions
      \end{enumerate}
      \end{minipage}
   }
   \BlankLine

   \Subst \(\leftarrow\) \SubstGen{\(M^{\text{new}}\), \(P^{\text{new}}(\varphi)\)}\;
   \Subst \(\leftarrow\) \Subst \(\cup\) \SubstGen{\(M^{\text{new}}\), \(P(\varphi)\)}\;
   \Subst \(\leftarrow\) \Subst \(\cup\) \SubstGen{\(M\), \(P^{\text{new}}(\varphi)\)}\;
   \BlankLine

@@

   \textcolor{ourblueviolet}{
   \MaxMono \(\leftarrow\) \Max{\Min{\MonoFloor, \(|M\cup M^{\text{new}}| \cdot \)\MonoMult}, \MonoCap}\;
   \MaxPoly \(\leftarrow\) \Max{\Min{\PolyFloor, \(|P(\varphi)\cup P^{\text{new}}(\varphi)| \cdot \)\PolyMult}, \PolyCap}\;
   }

@@

   \UsedSubst \(\leftarrow\emptyset\)\;
   \ForEach{\(\sigma\in \Subst\)}{
      \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in P(\varphi)\)}{
            \eIf{\TyVars{\((\upsilon_1, \dots, \upsilon_n)\)} \(\subseteq\) \Domain{\(\sigma\)}}{
               \textcolor{ourblueviolet}{
               \uIf{\(|M^{\text{next}}| < \MaxMono\) \label{algo:relative_bound}}{
                  \textcolor{black}{
                  \(M^{\text{next}}(f)\leftarrow M^{\text{next}}(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
                  \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;}
               }
               \uElseIf{\(|P^{\text{next}}(\varphi)| \geq \MaxPoly\) }{
                  blablabliblou\;
                  \(M^\text{next} \leftarrow M^{\text{next}}\setminus (M^\text{old} \cup M^\text{new})\)\;
                  \(P^\text{next}(\varphi)\leftarrow P^{\text{next}}(\varphi)\setminus (P^\text{old}(\varphi)\cup P^\text{new}(\varphi))\)\;
                  \Return \((M^\text{next}, P^\text{next}(\varphi), \UsedSubst)\)
               }}
            }
            {
               \textcolor{ourblueviolet}{
               \If{\(|P^{\text{next}}(\varphi)| < \MaxPoly\)\label{algo:ref_2}} {
               \textcolor{black}{
               \(P^{\text{next}}(\varphi)(f)\leftarrow P^{\text{next}}(\varphi)(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
               \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;
            }}
            }

@@

   \Return \((M^{\text{next}}\setminus (M^\text{old} \cup M^\text{new}), P^{\text{next}}(\varphi)\setminus (P^\text{old}(\varphi)\cup P^\text{new}(\varphi)), \UsedSubst)\)
}

\end{algorithm}
\end{quote}
\caption{Formula Monomorphisation Step}
\label{mono_step}
\end{figure}

@@

\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\SetKwFunction{SubstGen}{substitution\_generation}
\Fn(){\SubstGen{\(M, P(\varphi)\)}}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}
   \SetKwData{SubstLimit}{substitution\_limit}
   \SetKwFunction{Success}{Success}\SetKwFunction{Compatible}{compatible}
   \SetKwFunction{Match}{match}
   \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
   \SetKw{Break}{break}

   \KwData{%
      \hspace{-2.7em}\begin{minipage}[t]{.8\textwidth}
      \begin{enumerate}
         \item[] monomorphic type argument map \(M\)
         \item[] non-monomorphic type argument map \(P(\varphi)\)
      \end{enumerate}
      \end{minipage}
   }
   \KwResult{set of substitutions}
   \BlankLine

@@

   \Fn(){\MonoSubst{\(S^{\text{all}}, S^\text{acc}, \sigma^\text{acc}, V\)}}{

   \KwData{%
      \hspace{-2.7em}\begin{minipage}[t]{.8\textwidth}
      \begin{enumerate}
         \item[] set of substitutions \(S^{\text{all}}\) and \(S^\text{acc}\)
         \item[] single substitution \(\sigma^\text{acc}\)
         \item[] set of type variables \(V\)
      \end{enumerate}
      \end{minipage}
   }

@@

   \eIf{\(V = \emptyset\)}{
    \Return \(S^\text{acc}\cup\{\sigma^\text{acc}\}\)\;%}
   }
   {
      \Let \(\alpha \) such that \(\alpha\in V\)\;
      \ForEach{\(\sigma^\text{next}\in S^{\text{all}}\) such that, \(\alpha\in\Domain{\(\sigma^\text{next}\)}\) \And \(\sigma^\text{acc}\) compatible with \(\sigma^\text{next}\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|S^\text{acc}|<\Limit\)}{
            \textcolor{black}{
            \(S^\text{acc}\leftarrow \MonoSubst{\(S^{\text{all}}, S^\text{acc}, \sigma^\text{next}\circ\sigma^\text{acc}, V\setminus\Domain{\(\sigma^\text{next}\)}\)}\)\;}
         }
         {\Return \(S^\text{acc}\)}}
      }
      \Return \(S^\text{acc}\)\;
   }
   }

\end{algorithm}
\end{quote}
\caption{Monomorphising Substitutions}
\label{mono_substs}
\end{figure}

@@

To monomorphise a polymorphic formula, we first compute its \emph{monomorphising substitutions} (\ref{mono_substs}). A substitution \(\sigma\) is a monomorphising substitution for a formula \(\varphi\) if \(\sigma(\varphi)\) is monomorphic. \ref{mono_substs} is a recursive function, given a set of type variables \(V\) and a pool of substitutions \(S^\text{all}\), it will select a substitution \(\sigma^\text{next}\) from \(S^\text{all}\) which instantiates at least one of the type variables in \(V\). It is important that \(\sigma^\text{next}\) be compatible with \(\sigma^\text{acc}\) so that they may be composed and the function recursively called to instantiate the remaining type variables.

>>>>

\item[\labelitemi] Given a formula \(\varphi \in \Phi\) and a symbol \(f\) occurring in \(\varphi\), the set \(P(\varphi)(f)\) consists of all tuples of type arguments to which \(f\) is applied in \(\varphi\) and which contains a type variable. For example, if \(\sym{foldl}\langle \ty{nat}, \ty{list}(\alpha)\rangle\) occurs in \(\varphi\), then \((\ty{nat}, \ty{list}(\alpha)) \in P(\varphi)(\sym{foldl}) \).
\end{enumerate}

@@

   \item For each formula \(\varphi \in \Phi\) and for each symbol \(f\) occurring in \(\varphi\):
   \begin{enumerate}
    \item[2.1.] For each tuple \((\tau_1, \dots, \tau_n) \in  M(f)\)        and each tuple \((\upsilon_1, \dots, \upsilon_n) \in P(\varphi)(f)\),
     for each \(i\), match \(\upsilon_i\) against \(\tau_i\), yielding the substitution \(\sigma_i\) in case of success.

@@

\SetKwFunction{IterMono}{iterative\_monomorphisation}
\begin{figure}[b]
\begin{quote}
\begin{algorithm}[H]
\Fn(){\(\IterMono(\Phi)\)}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}\SetKw{To}{to}

   \SetKwData{Loop}{num\_loops}\SetKwData{NewFormulae}{\(\Psi\)}
   \SetKwData{Problem}{\(\Phi\)}\SetKwData{AllSubst}{\(S\)}
   \SetKwData{Limit}{total\_new\_formulae}

   \SetKwFunction{Init}{initialisation}\SetKwFunction{FStep}{formula\_mono\_step} 
   \SetKwFunction{TyVars}{type\_variables}\SetKwFunction{MonoSubst}{mono\_substs}
   \SetKwFunction{GenFormulae}{generate\_mono\_formulae}

   \KwData{finite set \(\Phi\) of polymorphic formulae}
   \KwResult{finite set of monomorphic formulae}

   \((M_{\text{old}}, P_{\text{old}}) \leftarrow (\emptyset, \emptyset)\)\;
   \((M_{\text{new}}, P_{\text{new}}) \leftarrow \Init(\Problem)\)\;

@@

   \For{\(i = 1\) \To \textcolor{ourblueviolet}{\Loop}}{
      \((M_{\text{next}}, P_{\text{next}})\leftarrow(\emptyset,\emptyset)\)\;
      \ForEach{\(\varphi \in \Problem\)} {
         \((M_{\Delta}, P_{\Delta}(\varphi), S_{\Delta})\leftarrow \FStep(M_{\text{old}}, P_{\text{old}}(\varphi), M_{\text{new}}, P_{\text{new}}(\varphi))\)\;
         \(\AllSubst\leftarrow\AllSubst\cup S_{\Delta}\)\;
         \(M_{\text{next}}\leftarrow M_{\text{next}}\cup M_{\Delta}\)\;
         \(P_{\text{next}}(\varphi)\leftarrow P_{\text{next}}(\varphi)\cup P_{\Delta}(\varphi)\)\;
      }

      \((M_{\text{old}}, M_{\text{new}})\leftarrow (M_{\text{old}}\cup M_{\text{new}}, M_{\text{next}})\)\;
      \((P_{\text{old}}, P_{\text{new}})\leftarrow (P_{\text{old}}\cup P_{\text{new}}, P_{\text{next}})\)\;

@@

   \Return \(\GenFormulae(\Phi, S)\)
}
\end{algorithm}
\end{quote}
\caption{Pseudocode for iterative monomorphisation}
\label{iter_mono}
\end{figure}

\SetKwFunction{GenFormulae}{generate\_mono\_formulae}
\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\Fn(){\(\GenFormulae(\Phi, S)\)}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}\SetKw{To}{to}

   \SetKwData{Loop}{num\_loops}\SetKwData{NewFormulae}{\(\Psi\)}
   \SetKwData{Problem}{\(\Phi\)}\SetKwData{AllSubst}{\(S\)}
   \SetKwData{Limit}{total\_new\_formulae}

   \SetKwFunction{Init}{initialisation}\SetKwFunction{FStep}{formula\_mono\_step} 
   \SetKwFunction{TyVars}{type\_variables}\SetKwFunction{MonoSubst}{mono\_substs}
   \SetKwFunction{GenFormulae}{generate\_formulae}

   \KwData{\begin{minipage}[t]{.8\textwidth}
     \strut finite set \(\Phi\) of polymorphic formulae \\
     finite set $S$ of substitutions\strut
     \end{minipage}}
   \KwResult{finite set of monomorphic formulae}

   \NewFormulae \(\leftarrow\emptyset\)\;

   \ForEach{\(\varphi\in\Phi\) such that $\varphi$ is non-monomorphic}{
      \ForEach{\(\sigma\in \MonoSubst{\AllSubst, \(\TyVars{\(\varphi\)}\), \(\emptyset\), \(\{\}\)}\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|\NewFormulae|<\Limit\)}{
            \textcolor{black}{
            \(\NewFormulae\leftarrow\NewFormulae\cup \{\sigma(\varphi)\}\)\;}
         }
         {\Return \NewFormulae}
      }}
   }

   \Return \NewFormulae
}
\end{algorithm}
\end{quote}
\caption{Pseudocode for monomorphic formula generation}
\label{genformulae}
\end{figure}

@@

   %\LinesNumbered
   \SetKwFunction{SubstGen}{matches}\SetKwFunction{TyVars}{type\_variables}
   \SetKwFunction{FMonoStep}{formula\_mono\_step}

   \Fn(){\FMonoStep{\(M_{\text{old}}, M_{\text{new}}, P_{\text{old}}(\varphi), P_{\text{new}}(\varphi)\)}}{
   \SetKwData{TypeVars}{type\_variables}
   \SetKwData{MaxMono}{max\_mono\_args}
   \SetKwData{MaxPoly}{max\_poly\_args}
   \SetKwData{UsedSubst}{\(S\)}

@@

   \KwData{\begin{minipage}[t]{.8\textwidth}
     \strut old and new monomorphic type argument maps \(M_{\text{old}}, M_{\text{new}}\) \\
     old and new non-monomorphic type argument maps \(P_{\text{old}}(\varphi), P_{\text{new}}(\varphi)\)\strut
      \end{minipage}
   }
   \KwResult{\begin{minipage}[t]{.8\textwidth}
      \strut monomorphic type argument map \\
      non-monomorphic type argument map \\
      set of substitutions\strut
      \end{minipage}
   }
   \BlankLine

@@

   \textcolor{ourblueviolet}{
   \MaxMono \(\leftarrow\) \Max{\Min{\MonoFloor, \(|M_{\text{old}}\cup M_{\text{new}}| \cdot \MonoMult\)}, \MonoCap}\;
   \MaxPoly \(\leftarrow\) \Max{\Min{\PolyFloor, \(|P_{\text{old}}(\varphi)\cup P_{\text{new}}(\varphi)| \cdot \PolyMult\)}, \PolyCap}\;
   }

@@

   \UsedSubst \(\leftarrow\emptyset\)\;

   \ForEach{\(\sigma\in {}\)\SubstGen{\(M_{\text{new}}\), \(P_{\text{new}}(\varphi)\)}
\(\cup\) \SubstGen{\(M_{\text{new}}\), \(P_{\text{old}}(\varphi)\)}
\(\cup\) \SubstGen{\(M_{\text{old}}\), \(P_{\text{new}}(\varphi)\)}}{
      \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in P_{\text{old}}(\varphi)\)}{
            \eIf{\TyVars{\((\upsilon_1, \dots, \upsilon_n)\)} \(\subseteq\) \Domain{\(\sigma\)}}{
               \textcolor{ourblueviolet}{
               \uIf{\(|M_{\text{next}}| < \MaxMono\) \label{algo:relative_bound}}{
                  \textcolor{black}{
                  \(M_{\text{next}}(f)\leftarrow M_{\text{next}}(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
                  \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;}
               }
               \uElseIf{\(|P_{\text{next}}(\varphi)| \geq \MaxPoly\) }{
                  \(M_{\text{next}} \leftarrow M_{\text{next}}\setminus (M_{\text{old}} \cup M_{\text{new}})\)\;
                  \(P_{\text{next}}(\varphi)\leftarrow P_{\text{next}}(\varphi)\setminus (P_{\text{old}}(\varphi)\cup P_{\text{new}}(\varphi))\)\;
                  \Return \((M_{\text{next}}, P_{\text{next}}(\varphi), \UsedSubst)\)
               }}
            }
            {
               \textcolor{ourblueviolet}{
               \If{\(|P_{\text{next}}(\varphi)| < \MaxPoly\)\label{algo:ref_2}} {
               \textcolor{black}{
               \(P_{\text{next}}(\varphi)(f)\leftarrow P_{\text{next}}(\varphi)(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
               \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;
            }}
            }

@@

   \Return \((M_{\text{next}}\setminus (M_{\text{old}} \cup M_{\text{new}}), P_{\text{next}}(\varphi)\setminus (P_{\text{old}}(\varphi)\cup P_{\text{new}}(\varphi)), \UsedSubst)\)
}

\end{algorithm}
\end{quote}
\caption{Pseudocode for formula monomorphisation step}
\label{mono_step}
\end{figure}

@@

\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\SetKwFunction{SubstGen}{matches}
\Fn(){\SubstGen{\(M, P(\varphi)\)}}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}
   \SetKwData{SubstLimit}{substitution\_limit}
   \SetKwFunction{Success}{Success}\SetKwFunction{Compatible}{compatible}
   \SetKwFunction{Match}{match}
   \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
   \SetKw{Break}{break}

   \KwData{\begin{minipage}[t]{.8\textwidth}
      \strut monomorphic type argument map \(M\) \\
      non-monomorphic type argument map \(P(\varphi)\)\strut
      \end{minipage}
   }
   \KwResult{set of substitutions}
   \BlankLine

@@

   \Fn(){\MonoSubst{\(S_{\text{all}}, S^\text{acc}, \sigma^\text{acc}, V\)}}{

   \KwData{%
      \hspace{-2.7em}\begin{minipage}[t]{.8\textwidth}
      \begin{enumerate}
         \item[] set of substitutions \(S_{\text{all}}\) and \(S^\text{acc}\)
         \item[] single substitution \(\sigma^\text{acc}\)
         \item[] set of type variables \(V\)
      \end{enumerate}
      \end{minipage}
   }

@@

   \eIf{\(V = \emptyset\)}{
    \Return \(S^\text{acc}\cup\{\sigma^\text{acc}\}\)\;%}
   }
   {
      \Let \(\alpha \) such that \(\alpha\in V\)\;
      \ForEach{\(\sigma_{\text{next}}\in S_{\text{all}}\) such that, \(\alpha\in\Domain{\(\sigma_{\text{next}}\)}\) \And \(\sigma^\text{acc}\) compatible with \(\sigma_{\text{next}}\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|S^\text{acc}|<\Limit\)}{
            \textcolor{black}{
            \(S^\text{acc}\leftarrow \MonoSubst{\(S_{\text{all}}, S^\text{acc}, \sigma_{\text{next}}\circ\sigma^\text{acc}, V\setminus\Domain{\(\sigma_{\text{next}}\)}\)}\)\;}
         }
         {\Return \(S^\text{acc}\)}}
      }
      \Return \(S^\text{acc}\)\;
   }
   }

\end{algorithm}
\end{quote}
\caption{Pseudocode for monomorphising substitution generation}
\label{mono_substs}
\end{figure}

@@

To monomorphise a polymorphic formula, we first compute its \emph{monomorphising substitutions} (\ref{mono_substs}). A substitution \(\sigma\) is a monomorphising substitution for a formula \(\varphi\) if \(\sigma(\varphi)\) is monomorphic. \ref{mono_substs} is a recursive function, given a set of type variables \(V\) and a pool of substitutions \(S^\text{all}\), it will select a substitution \(\sigma_{\text{next}}\) from \(S^\text{all}\) which instantiates at least one of the type variables in \(V\). It is important that \(\sigma_{\text{next}}\) be compatible with \(\sigma^\text{acc}\) so that they may be composed and the function recursively called to instantiate the remaining type variables.
